### 2017-02-23 2017mergingmerges/section/XX-Related.tex
temporary notes about synchronous languages and regular dataflow

>>>>

\subsection{Scratch: dataflow}
Synchronous languages (LUSTRE, Lucy-n, SIGNAL) view streams as -- abstractly -- functions from time to values.
(Or time to maybe values, if they allow holes).
This forces them into an inherently \emph{push} based system, but what is pushing here is the procession of time.
If you have two streams, you cannot pull from one and the other: when time advances, it advances for all streams.
Thus, synchronous languages cannot encode value-dependent pulling - nor can they encode append.
Append could be seen as a special kind of value-dependent access pattern, if you treat the end of the stream as a special kind of value.

Non-synchronous streaming languages, such as Lucid, may allow some kind of value-dependent pulling, because they do not enforce causality.
However, they rely on dynamic scheduling.

Do not to confuse regular dataflow, also known as \emph{synchronous data flow}, with synchronous languages.
Regular dataflow graphs are dataflow graphs where each edge is annotated with the rate.

Dataflow languages:
\begin{itemize}
\item Lucid Synchrone
\item LUSTRE \cite{halbwachs1991synchronous}.
Dataflow language: causal analysis, clock calculus, synchronous.
Uses more or less syntactic equality to check equivalence of clocks.
\item Lucy-n: extension of LUSTRE with periodic clocks to ensure static scheduling \cite{mandel2010lucy}, \cite{benveniste2003synchronous}
\item SIGNAL \cite{le2003polychrony}
causal, clocks, synchronous
\item StreaMIT \cite{thies2002streamit}:
Regular/synchronous dataflow, ie statically known rates.
Only allows limited splits and joins: round robin and duplication for splits, round robin and combination for joins. 
Does not support fully general graphs - instead using combinators to introduce a (split/join) and a combinator for a feedback loop.
\item Prelude: extension of LUSTRE, seems very similar to Lucy-n \cite{pagetti2011multi}
\item Ptolemy (II)
\end{itemize}

Models:
\begin{itemize}
\item Synchronous dataflow (SDF):
all rates are statically known,
\item Cyclo-static dataflow (CSDF) where the number of produced/consumed tokens varies periodically, sounds very similar to Lucy-n's periodic clocks \cite{mandel2010lucy}
\item Boolean dataflow (BDF), generally undecidable to statically schedule, so must fall back to dynamic scheduling \cite{buck1993scheduling}
\item Integer dataflow (IDF), (also falls back to dynamic) \cite{buck1994static}
Clustering to find control structure: tries to recover structures like ifs, cases and loops from the dataflow graph.
Extends previous BDF \cite{buck1993scheduling} with loops that repeat some number of times, and multi-way ifs.
Very rigid and only supports limited control flow structures.
Unclear how a merge or append would be clustered by this system.
\item Parameterized dataflow (PDF),  \cite{bhattacharya2001parameterized}
\item Schedulable parametric dataflow (SPDF),  \cite{fradet2012spdf}
\end{itemize}

Dataflow languages tend to fall back on less performant dynamic scheduling, when programs cannot be statically scheduled.
For example \citet{bouakaz2013real} uses static scheduling for special cases like cyclo-static dataflow graphs, but otherwise uses dynamic earliest deadline first (EDF) scheduling.

When dataflow languages talk about a `merge' operator, they are talking about \emph{non-deterministic merge}. These are not Kahn processes.
